Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-27 paper/main.tex
reorganize subsections; only mention heterogeneity in applications as paragraph

By utilizing the provided data extraction methods, we can infer properties of the bacteria that aid
us in constructing a suitable computational model.
However, the construction of any mechanistic model also requires prior konwledge which can be used
to inform our analysis methods.
For example, by extracing the positions of bacteria across a time-series of microscopic images, we
can also calculate their lengths at these time-points and thus determine their rate of growth and
type of growth (see section~\ref{subsec:parameter-estimation-individual-treatment}), thus
effectively guiding our modeling steps.
As we will show later, extracted data in the form of cellular positions or segmentation masks can
also be used for comparing computational results with biological ground truth.

@@

We have already suspected (see Table~\ref{table:simulation-aspects}) that in order to properly
estimate parameters of our system, we need to account for natural cell-variability, thus assigning
parameters on a per-cell basis.

@@

\label{subsec:parameter-estimation-individual-treatment}
\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.28\textwidth]
        {figures/crm_fit-data-progression.png}
        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-distribution.pdf}%
        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \caption{
        \textbf{TODO Fix caption!}\\
        \textbf{Link to supplement for discussion of bar plot}\\
        % TODO use same y-axis across the plots
        (B) The subfigure shows the time-evolution of the rod-lengths for every individual
        bacterium.
        These growth curves are fitted individually with delayed exponential function.
        (C) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
        fitted with an exponential function.
        The uncertainty band is the standard deviation of the averaged datapoints.
        (C) Blue dots show the combination of estimated growth rates and delays for individual
        bacteria while the orange dot corresponds to the average.
        Ellipses around the datapoints are taken from the uncertainty of the fits.
        Time is given in units of frames of the movie from which the data have been extracted.
    }
    \label{fig:estimated-growth-rates}
\end{figure}

>>>>

By utilizing the provided data extraction methods, we can infer properties of the bacteria that aid
us in constructing a suitable computational model.
However, the construction of any mechanistic model also requires prior konwledge which can be used
to inform our analysis methods.
For example, by extracing the positions of bacteria across a time-series of microscopic images, we
can also calculate their lengths at these time-points and thus determine their rate of growth and
type of growth (see section~\ref{paragraph:parameter-estimation-individual-treatment}), thus
effectively guiding our modeling steps.
As we will show later, extracted data in the form of cellular positions or segmentation masks can
also be used for comparing computational results with biological ground truth.

@@

\paragraph{Cell-Cell Variability within the Dataset}
\label{paragraph:parameter-estimation-individual-treatment}
We have already discussed (see Section~\ref{sec:role-of-heterogeneity}) that in order to properly
estimate parameters of our system, we need to account for natural cell-variability, thus assigning
selected parameters on a per-cell basis.

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.28\textwidth]
        {figures/crm_fit-data-progression.png}
        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-distribution.pdf}%
        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \caption{
        \textbf{TODO Fix caption!}\\
        \textbf{Link to supplement for discussion of bar plot}\\
        % TODO use same y-axis across the plots
        (B) The subfigure shows the time-evolution of the rod-lengths for every individual
        bacterium.
        These growth curves are fitted individually with delayed exponential function.
        (C) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
        fitted with an exponential function.
        The uncertainty band is the standard deviation of the averaged datapoints.
        (C) Blue dots show the combination of estimated growth rates and delays for individual
        bacteria while the orange dot corresponds to the average.
        Ellipses around the datapoints are taken from the uncertainty of the fits.
        Time is given in units of frames of the movie from which the data have been extracted.
    }
    \label{fig:estimated-growth-rates}
\end{figure}
